% vim:set ft=tex:
\section{Load and Balance}
\label{design:load}

\newacronym{mpc}{MPC}{\gls{llc} misses per cycle}

What is load?
And depending on the answer to this question: What is balance?

As I describe in \ref{state:cfs}, \gls{cfs} defines a core's load as the amount of
execution time used by threads.
However, a core's load can also be defined as
the cache and memory usage, the sum of \gls{instpc}, or simply as the number of
assigned threads.
It is also imaginable to use a mix of the above as load measure.

The answer to the question: ``What is balance?" depends on the definition of
load.
Assuming the \gls{cfs} definition, a domain is balanced if the groups within
the domain carry an even amount of load.
If this assumption holds for each level in the domain hierarchy, the system is
in balance.
However, load only represents consumed computation time, thus, it balances time
and disregards space completely.

The opposite approach measures only \gls{mpc} and ignores computation
resources.
So the load of a domain is the sum of \gls{mpc} of all assigned threads.
A system is in balance, if each group within a domain has an equal number of
\gls{mpc}, or in a different formulation:
if no thread-to-core assignment reduces the imbalances between cores.
However, this metric suffers because of its disregard for remaining execution time.
While memory intense threads may perform better, computation intense threads
suffer due to migration and possibly thread count imbalances between
cores.
For example, if one thread produces a high number of \gls{mpc} it is assigned a
core exclusively, to meet its cache requirements.
If four other threads are computation intense and the sum of their \gls{mpc} is
still lower than the high \gls{mpc} threads, they will be sharing one core.
This system is in balance regarding \gls{mpc}, but regarding execution resource
it is perceived as heavily imbalanced.

A third metric is \gls{instpc}, which similar to load is a temporal balancing
metric.
It shows, how much instructions a thread was able to perform during its
run-time.
Similar to \gls{mpc}, this number is different for each thread.
CPU bound workloads tend to have a high number of \gls{instpc},
whereas memory bound workloads tend to have a much lower number, due to cache
misses and waiting for memory.
Also, instructions are not equally costly, which leads
\citeauthor{snavely_symbiotic_2000} (\cite{snavely_symbiotic_2000}) to their
weighted instruction measure.
However, at run-time it is not possible to pinpoint the weight of the current
instruction.

Balancing only by \gls{instpc} will lead to a similarly unfair distribution of
threads to cores as balancing by \gls{mpc}.
The disregard for cache usage will likely decrease the number of instructions
a cache sensitive thread is able to execute, as more cache misses are
generated.

The number of threads assigned to a core also impacts the progress of other
assigned threads.
However, only if the newly assigned thread is really executing and not a service,
which mostly idles and executes only on request.
Thread count ignores load, \gls{instpc}, and \gls{mpc} and is therefore
not of much use.

A mix of execution time, \gls{instpc}, and \gls{mpc} looks promising.
The main drawback of execution time is its obliviousness of cache usage.
Extending the execution time metric with \gls{mpc} knowledge should alleviate this.
A second approach can combine \gls{instpc} and \gls{mpc} metrics: I assume
\gls{instpc} and \gls{mpc} to be two orthogonal measures, hence I expect running a
high \gls{mpc} thread on the same core as a high \gls{instpc} thread to perform
well.
\\

% todo merit flaw table as summary

After discussing load and balance, it is also necessary to decide on a
threshold, before actively migrating threads.
The cost of a migration must not necessarily be paid to have a perfectly
balanced system.
Slight imbalances decrease the performance compared to a perfectly balanced
system, but as long as this decrease is less than the cost of a migration,
it is preferable to tolerate the imbalance.

Cache-heavy threads, determined by last-level cache misses per cycle, can have
two reasons for a high miss count: first, they reuse little memory and use new
memory in a not prefetchable manner, otherwise the prefetcher would not miss;
second, the cache working set is large and parts of the working set were
evicted during the run-time of other programs.
Migrating the former kind does not hurt the program much, migrating the latter
kind, however, is costly as the complete cache working set must be reloaded
into the new core's cache.

While it is generally a bad idea to migrate high \gls{mpc} threads, a
clustering of high \gls{mpc} threads is worse and must trigger a migration of a
one of the high \gls{mpc} threads.

In conclusion, balancing should take place in time and in space.
While balancing in space takes priority over balancing in time, temporal
balancing evens out imbalances originating from spatial balancing.
The spatial balancing should act, if high imbalances regarding accumulated
\gls{mpc} per core arise, or if the difference of \gls{mpc} heavy threads per
core rises above one or two high \gls{mpc} threads per core.

Regarding temporal imbalances, the questions are: is there a thread that would
even out imbalances if migrated?
And: At which load differences does the load balancer start searching for such
a thread?
In the evaluation section I search for a good threshold value.
