% vim:set ft=tex:
\pagebreak
\section{Task Configuration}
\label{impl:config}

In sections \ref{design:isolation} and \ref{design:comm}, I discussed
task properties relevant to a load balancer.
The result of these discussions was to provide the load balancer with a static
configuration of a task's properties.
The following section shows how the configuration is used and implemented.

\paragraph{Configuration Parameters}
Listing \ref{impl:config:parameters} shows the implemented parameters.

\begin{lstlisting}[language={[5.2]Lua}, caption={Implemented set of
configuration parameters.},label={impl:config:parameters}]
min_prio  = <number>
max_prio  = <number>
sec       = <name>
rt        = <name>
distr     = <name>
clsvr     = <name>_{C|S}
\end{lstlisting}

The system designer needs not to define orthogonal parameters for
a task at the same time.
Either a task is in a security or in a real-time group, never in both.
The same goes for \texttt{clsvr\_C} and \texttt{clsvr\_S}:
A task cannot be client and server in the same group.
Also, \texttt{distr} and \texttt{clsvr} may not be combined, as the
\texttt{distr} property is assumed if a client task runs more than one thread.
Additionally, the system designer may not define more real-time and security
groups than physical cores available.
As mentioned in \ref{design:isolation} a real-time group could have more than
one core assigned, but this is currently not implemented.
In the future the name could be followed by a core count, to state the number
of cores necessary.
Then a combination of \texttt{clsvr} and \texttt{rt} makes sense, which is not
the case at the current state.

To configure the task a create call to the factory service of the load
balancer must be issued, containing the scheduler protocol as first parameter
and then a comma separated list of strings stating the configuration
parameters.
The factory will then return a capability on a channel to its proxy scheduler
encapsulating the task configuration.
\begin{comment}
Listing \ref{impl:config:proxy} shows an example for the described create call.

\begin{lstlisting}[language={[5.2]Lua},caption={Create call to the load
balancer factory containing protocol, priority band, and a client-server
group.},label={impl:config:proxy}]
tmFactory:create(
  L4.Proto.Scheduler,
  "min_prio = 0",
  "max_prio = 5",
  "clsvr = A_S")
\end{lstlisting}
\end{comment}

The capability received from the above create call replaces replaces the
inherited scheduler in the task's environment.
As the scheduler interface has not changed, the task accesses the scheduler via
\texttt{L4Re::Env::env()->scheduler()}.
Listing \ref{impl:config:taskcreation} displays a complete task creation in
the Ned configuration script.

\begin{lstlisting}[language={[5.2]Lua},caption={Task creation with replaced
scheduler in Ned startup script.},label={impl:config:taskcreation}]
L4.default_loader:start(
  {
    caps = {},
    scheduler = tmFactory:create(
      L4.Proto.Scheduler,
      "min_prio = 0", "max_prio = 5", "distr = A")
  },
  "rom/app_matrixmul"
)
\end{lstlisting}

Therefore, the only change necessary to the application are the configuration
parameters, which are encapsulated in the Ned startup script, shown in listing
\ref{config:ned_full}.
The load balancing service is completely transparent to the application.

\lstinputlisting[language={[5.2]Lua}, firstline=3, caption={Complete Benchmark
configuration}, label={config:ned_full}]{app_restart_mmul_parallel.cfg}


\paragraph{Configuration Groups}

The configuration arguments stated in the startup script are parsed by the
factory service and stored in a \texttt{config\_t} type
(listing \ref{config:config_t}).
The parser strips all white spaces from the configuration string and splits it
in two parts: before and after the equality sign.
The front part evaluates to priority or group type, whereas the rear part
evaluates to either an integer or string.
%The front part evaluates to minimum or maximum priority, or group type and the
%rear part evaluates to either an integer or a string identifier for the group.
If the group type is a client-server type, the last character is matched
against \texttt{C} or \texttt{S} to recognize client and server tasks.

Note, this configuration is per task, hence all threads of this task possess
the same configuration.
A configuration group must not necessarily span across several tasks.
The \texttt{distr} parameter, for example, can group all threads of one task
and ensures, that all threads of this task will be assigned to the maximal
number of available cores.
\\
\newline

\lstinputlisting[language=c++, caption={Configuration Type},
label={config:config_t}]{codeSamples/config.cc}
